---
title: "Formatting, alpha, and beta"
author: "Stefano Romano"
date: "Oct 2020"
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

# Load the needed packages
```{r, message=F}
list.of.packages <- c("ggplot2", "reshape2", "dplyr", "ggnewscale", "ggrepel", "lme4", "FSA",
                      "devtools", "ggpubr", "vegan", "RColorBrewer", "emmeans", "car")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

if (!requireNamespace("BiocManager", quietly = TRUE)){
  install.packages("BiocManager")}


list.of.bioc <- c("phyloseq", "grDevices", "microbiome")
new.packages <- list.of.bioc[!(list.of.bioc %in% installed.packages()[,"Package"])]
if(length(new.packages)) BiocManager::install(new.packages)

library(dplyr)
library(lme4)
library(RColorBrewer)
library(phyloseq)
library(microbiome)
library(ggpubr)
library(ggplot2)
library(reshape2)
library(grDevices)
library(vegan)
library(devtools)
library(ggrepel)
library(ggnewscale)
library(FSA)
library(emmeans)
library(car)

# To install pairwiseAdonis visit: https://github.com/pmartinezarbizu/pairwiseAdonis
library(pairwiseAdonis)
# To install ggConvexHull visit: https://github.com/cmartin/ggConvexHull
library(ggConvexHull)
```


# Load the species rel ab. calculated including the unknown.

We are using only the species.
Note, this does not include mice and virus sequences.

```{R, laod data and split taxonomic ranks}
tax <- read.delim("Raw_data/merged_metaphlan3_spec.txt", sep = "\t", stringsAsFactors = F)
```

### Load meta data and format

```{r, message=F}
# Load metadata 
m<-read.delim("Metadata.txt", header = T, stringsAsFactors = F)
```

### Double check that datset is clean

```{R}
grep("Eukar*", rownames(sp))
grep("Chlorop*", rownames(sp))
grep("Mitoc*", rownames(sp))
# Dataset is clean

sapply(m, class)
# Need to make them factors
m$receiver<-factor(m$receiver, levels = c("Y", "O", "M"))
m$treat<-factor(m$treat, levels = c("Y", "O", "M", "Abx"))
m$timegroups<-as.factor(m$timegroups)
m$Timepoint<-factor(m$Timepoint, levels = c("PRE_Abx", "POST_FMT", "POST_ABX", "END"))
m$cage_tg<-as.factor(m$cage_tg)
m$groups<-as.factor(m$groups)

sapply(m, class)
```


Create the phyloseq object
```{r, message=FALSE, warning=FALSE}
sp_phylo<-phyloseq(otu_table(sp, taxa_are_rows = T), sample_data(m))
sp_phylo
```

# Alpha

Calculate number of observed species.
```{r}
rich<-richness(sp_phylo, index = "observed")
dir.create("Taxonomy/Figures", recursive = T)
```

Merge the data with meta-data and and plot.
```{R}
all(rownames(rich) == rownames(meta(sp_phylo)))
rich_m<-cbind.data.frame(rich, meta(sp_phylo))
rich_m$Timepoint<-factor(as.character(rich_m$Timepoint), levels = c("PRE_Abx", "POST_ABX", "POST_FMT", "END"))

ggplot(rich_m, aes(Timepoint, observed, fill = Timepoint)) +
  geom_boxplot(outlier.colour = "none") +
  geom_jitter(width = 0.2, aes(color = cage)) +
  facet_wrap(~groups) +
  scale_fill_brewer() +
  theme_bw() +
  labs(title = "Number of observed species per mice type and timepoint")+
  theme(axis.text.x = element_text(angle = 90))

ggsave("Taxonomy/Figures/Observed_species.svg", device = "svg", width = 8)

ggplot(rich_m, aes(Timepoint, observed, fill = Timepoint)) +
  geom_boxplot(outlier.colour = "none") +
  geom_jitter(width = 0.2) +
  facet_wrap(~receiver) +
  scale_fill_brewer() +
  theme_bw() +
  labs(title = "Number of observed species per mice type and timepoint")+
  theme(axis.text.x = element_text(angle = 90))

ggsave("Taxonomy/Figures/Observed_species_allmice.svg", device = "svg", width = 8, height = 5)

```

Plot the start for all groups.
```{r}

rich_m_s<-subset(rich_m, Timepoint == "PRE_Abx")
ggplot(rich_m_s, aes(receiver, observed, fill = receiver)) +
  geom_boxplot(outlier.colour = "none") +
  geom_jitter(width = 0.2, aes(shape = cage_mice), size = 4) +
  scale_fill_brewer() +
  scale_shape_manual(values = 1:32) +
  theme_bw() +
  labs(title = "Number of observed species per mice group Pre_Abx")+
  theme(axis.text = element_text(size = 12), legend.position = "none")
ggsave("Taxonomy/Figures/Observed_species_allmice_start.svg", device = "svg", width = 8, height = 5)

```

Verify whether there is a statistically significant difference between number of observed species.
```{r}
s_m<-subset(rich_m, Timepoint == "PRE_Abx")

x<-lmer(observed ~ receiver + (1|cage_mice), data = s_m)
Anova(x, type = "III")
emm_options(contrast = list(infer = c(TRUE, T)))
emmeans(x, pairwise ~ receiver, adjust = "fdr") 
```

Investigate by mice type: change in observed species across timepoints.
```{R}
# Y
s_m<-subset(rich_m, receiver == "Y")
x<-lmer(observed ~ Timepoint + (1|cage_mice), data = s_m)
emmeans(x, pairwise ~ Timepoint)

# O
s_m<-subset(rich_m, receiver == "O")
x<-lmer(observed ~ Timepoint + (1|cage_mice), data = s_m)
emmeans(x, pairwise ~ Timepoint)

# M
s_m<-subset(rich_m, receiver == "M")
x<-lmer(observed ~ Timepoint + (1|cage_mice), data = s_m)
emmeans(x, pairwise ~ Timepoint)
```

# Ordination

## Calculate beta diversity
Both JSD and Bray-Curtis have been used. 
The latter has been used as resulted in a clearer pattern in the ordination.

```{r}
sp_jsd<-phyloseq::distance(sp_phylo, method= "jsd", na.rm=F)
sp_bray<-phyloseq::distance(sp_phylo, method= "bray", na.rm=F)

# Make PCoA plot
sp_ord_jsd<-ordinate(sp_phylo, method = "PCoA", distance = sp_jsd)
sp_ord_bray<-ordinate(sp_phylo, method = "PCoA", distance = sp_bray)

```


To calculate the proportion of variance explained we need to account for the negative eigenvalues.
Apparently _phyloseq_ does not do that.
We will account for negative eigenvalues manually.
Negative eigenvalues will be set to 0

## JSD

```{r, message=FALSE, warning=FALSE}
eig<-sp_ord_jsd$values
eig[eig < 0]<-0
axis1<-paste("Axis 1 [", round(eig[1,1]/sum(eig[,1]),2)*100, "%]", sep = "")
axis2<-paste("Axis 2 [", round(eig[2,1]/sum(eig[,1]),2)*100, "%]", sep = "")

(jsd_pcoa<-plot_ordination(sp_phylo, sp_ord_jsd, 
                "samples", color="Timepoint", shape = "receiver") +
                 geom_point(aes(shape = receiver), color = "black",size = 4.5) +
                theme_bw() +
                scale_colour_brewer(palette = "Set2") +
                geom_point(size = 3.5) +
    labs(title = "Microbiota structure inferred using Jensen-Shannon divergence (JSD)", shape = "Receiver") +
    theme(plot.title = element_text(face = "bold"), panel.grid = element_blank()) +
    xlab(axis1) +
    ylab(axis2))

ggsave(jsd_pcoa, filename = "Taxonomy/Figures/PCoA_overall_jsd.svg", 
       device = "svg", width = 8, height = 8)

```

## BRAY

```{r, message=FALSE, warning=FALSE}
eig<-sp_ord_bray$values
eig[eig < 0]<-0
axis1<-paste("Axis 1 [", round(eig[1,1]/sum(eig[,1]),2)*100, "%]", sep = "")
axis2<-paste("Axis 2 [", round(eig[2,1]/sum(eig[,1]),2)*100, "%]", sep = "")

(bray_pcoa<-plot_ordination(sp_phylo, sp_ord_bray, 
                "samples", color="Timepoint", shape = "receiver") +
                 geom_point(aes(shape = receiver), color = "black",size = 4.5) +
                theme_bw() +
                scale_colour_brewer(palette = "Set2") +
                geom_point(size = 3.5) +
    labs(title = "Microbiota structure inferred using Bray-Curtis dissimilarities", shape = "Receiver") +
    theme(plot.title = element_text(face = "bold"), panel.grid = element_blank()) +
    xlab(axis1) +
    ylab(axis2))
ggsave(bray_pcoa, filename = "Taxonomy/Figures/PCoA_overall_bray.svg", 
       device = "svg", width = 8, height = 8)

```

## Permanova
We need to account for the cage effect in the dataset.
Format a df for permanova.
```{r, message=FALSE, warning=FALSE}
set.seed(987) 
all(rownames(t(otu_table(sp_phylo))) == rownames(meta(sp_phylo)))
df_m<-cbind.data.frame(t(otu_table(sp_phylo)),meta(sp_phylo))
```


#### JSD
```{r}
mod1 <- ordinate(sp_phylo, method = "CAP", formula = ~ Timepoint, Condition(cage_mice), distance = sp_jsd)
set.seed(56987)
h <- how(blocks = df_m$cage_mice, nperm = 10000)

sed_perm<-anova.cca(mod1, data = df_m, permutations = h, parallel = 2)
sed_perm
RsquareAdj(mod1)
```

#### Bray
```{r}
mod1 <- ordinate(sp_phylo, method = "CAP", formula = ~ Timepoint, Condition(cage_mice), distance = sp_bray)
set.seed(56987)
h <- how(blocks = df_m$cage_mice, nperm = 10000)

sed_perm<-anova.cca(mod1, data = df_m, permutations = h, parallel = 2)
sed_perm
RsquareAdj(mod1)
```

#### Pairwise permanova
The p-values seem to not be adjusted in the function _pairwise.adonis2_
Hence, they will be corrected manually.

#### JSD
```{r}
(adp_treat_timegroups_strata_cage<-pairwise.adonis2(sp_jsd ~ Timepoint, perm = 10000, data = df_m, strata = 'cage_mice'))
pv<-unlist(lapply(adp_treat_timegroups_strata_cage[-1], function(x) x$`Pr(>F)`[1]))
p.adjust(pv, method = "BH")
```

#### BRAY
```{r}
(adp_treat_timegroups_strata_cage<-pairwise.adonis2(sp_bray ~ Timepoint, perm = 10000, data = df_m, strata = 'cage_mice'))
pv<-unlist(lapply(adp_treat_timegroups_strata_cage[-1], function(x) x$`Pr(>F)`[1]))
p.adjust(pv, method = "BH")
```

## Test differences only in Old and Young Pre and Post FMT
```{r, message=FALSE, warning=FALSE}
pre_end_sp_phylo<-subset_samples(sp_phylo, Timepoint=="PRE_Abx" | Timepoint=="POST_FMT"  )
pre_end_sp_phylo<-subset_samples(pre_end_sp_phylo, receiver!="M" & treat !="M")
pre_end_sp_phylo<-subset_samples(pre_end_sp_phylo,  treat!= "Abx")

pre_end_sp_phylo<-filter_taxa(pre_end_sp_phylo, function(x) sum(x) > 0, TRUE)

# calculate JSD
pre_end_jsd<-phyloseq::distance(pre_end_sp_phylo, method= "jsd", na.rm=F)

df_m<-meta(pre_end_sp_phylo)
```

Perform pairwise comparisons.
```{r}
(pre_post_yo_oy_strata<-pairwise.adonis2(pre_end_jsd ~ timegroups,perm = 10000, data = df_m, strata = 'cage_mice'))
pv<-unlist(lapply(pre_post_yo_oy_strata[-1], function(x) x$`Pr(>F)`[1]))
p.adjust(pv, method = "BH")
```

# Ordination Pre and Post FMT

Function _bedadisp_ will be used.
PCoA will then be extracted to run _envfit_ on it.
```{r, message=F}
bd<-betadisper(pre_end_jsd, group = df_m$timegroups, add = T, type = "centroid")
plot(bd)
```

Make graph
```{r}
# Select 2 axis
ord_meta<-bd$vectors[,1:2]
all(rownames(df_m) == rownames(ord_meta))
ord_meta<-cbind.data.frame(ord_meta, df_m)

# Caluclate proportion of variacne explained 
eig<-bd$eig # There are no negative eigen
axis1<-paste("Axis 1 [", round(eig[1]/sum(eig),2)*100, "%]", sep = "")
axis2<-paste("Axis 2 [", round(eig[2]/sum(eig),2)*100, "%]", sep = "")

# Make a graph 
(g<-ggplot(ord_meta, aes(PCoA1, PCoA2, fill = Timepoint)) +
  geom_convexhull(alpha = 0.3, aes(fill = Timepoint, colour = timegroups), show.legend = F) +
  scale_color_manual(values = rep("black", 8)) +
  scale_fill_brewer(palette = "Set2") +
  new_scale_color() + 
  geom_point(color = "black", size = 4.5) +
  geom_point(aes(color = Timepoint), size = 3.5) +
  scale_color_brewer(palette = "Set2") +
  labs(title = "Gut microbiota structure pre and post FMT in young mice receiving aged stool", shape = "Receiver", color = "Donor") +
  xlab(axis1) +
  ylab(axis2) +
  theme_bw() +
  theme(plot.title = element_text(size = 14, face = "bold"), 
        panel.grid = element_blank(), 
        legend.position = "none"))
```

Plot ordination + Species that best correlate with the clustering.
```{r, message=FALSE, warning=FALSE}
sf <- envfit(bd, as.data.frame(t(otu_table(pre_end_sp_phylo))), 
             display = "sites", permutations = 10000) 

# Get the scores for plotting
vec.sp.df.scores<-as.data.frame(scores(sf, display = "vector"))
ef.df<-as.data.frame(sf$vectors$arrows*sqrt(sf$vectors$r))
ef.df$species<-rownames(ef.df)

# Colecting only taxa with significant pvalues
A <- as.list(sf$vectors)
pvals<-as.data.frame(A$pvals)
arrows<-as.data.frame(A$arrows*sqrt(A$r))
C<-cbind(arrows, pvals)
C$padj<-p.adjust(C$`A$pvals`, method = "fdr")
Cred<-subset(C,padj<0.005)
Cred <- cbind(Cred, Family = rownames(Cred))

# Collect only taxa that have the highest degree of variation along the axis
cap1<-subset(Cred, PCoA1 >= 0.43 | PCoA2 >= 0.43 | PCoA2 <= -0.43 |  PCoA1 <= -0.43)

taxa<-rownames(cap1)
shortaxa<-taxa
for(i in 1:length(shortaxa)){
  x<-strsplit(shortaxa[i], split = "|", fixed = T)[[1]][7]
  x<-gsub("s__", "", x)
  shortaxa[i]<-gsub("_", " ", x)

}

# Change naming and position manually 

cap1$shortaxa<-shortaxa

cap1_1<-subset(cap1, shortaxa == "Bifidobacterium animalis")
cap1_2<-subset(cap1, shortaxa == "Adlercreutzia equolifaciens")

cap1_4<-subset(cap1, shortaxa == "Asaccharobacter celatus")
cap1_19<-subset(cap1, shortaxa == "Firmicutes bacterium ASF500")
cap1_19$shortaxa<-"Firmicutes sp ASF500"
cap1_6<-subset(cap1, shortaxa == "Enterorhabdus caecimuris")
cap1_7<-subset(cap1, shortaxa == "Bacteroides caecimuris")
cap1_8<-subset(cap1, shortaxa == "Bacteroides uniformis")
cap1_5<-subset(cap1, shortaxa == "Muribaculaceae bacterium DSM 103720")
cap1_5$shortaxa<-"Muribaculaceae sp DSM 103720"
cap1_9<-subset(cap1, shortaxa == "Parabacteroides distasonis")
cap1_10<-subset(cap1, shortaxa == "Enterococcus faecalis")
cap1_11<-subset(cap1, shortaxa == "Lactobacillus johnsonii")
cap1_12<-subset(cap1, shortaxa == "Acutalibacter muris")
cap1_13<-subset(cap1, shortaxa == "Clostridium cocleatum")
cap1_14<-subset(cap1, shortaxa == "Prevotella sp MGM2")
cap1_15<-subset(cap1, shortaxa == "Faecalibaculum rodentium")
cap1_16<-subset(cap1, shortaxa == "Parabacteroides goldsteinii")
cap1_17<-subset(cap1, shortaxa == "Oscillibacter sp 1 3")
cap1_18<-subset(cap1, shortaxa == "Escherichia coli")
cap1_3<-subset(cap1, shortaxa == "Akkermansia muciniphila")
cap1_20<-subset(cap1, shortaxa == "Dorea sp 5 2")


library(ggrepel)
gs<- g +   geom_segment(data = cap1, stat = "identity",
              mapping = aes(x = 0, xend = PCoA1/3, y = 0, yend = PCoA2/3),
              arrow = arrow(length = unit(2, "mm"), type = "open"), 
              colour = "#C0C0C0", 
              inherit.aes = F, 
              linejoin = "mitre",
              arrow.fill = "#C0C0C0") +
  geom_label_repel(data = cap1_1,  #"Bifidobacterium animalis"
            stat = "identity",
            mapping = aes(x = PCoA1/3, y = PCoA2/3, label=cap1_1$shortaxa),
            segment.size  = 0.5,
            segment.color = "#A9A9A9",
            direction = "both",
            size = 4, nudge_x = 0.5, nudge_y = 0.01, 
            inherit.aes = F, 
            point.padding = 0.5,
            seed = 987, max.iter = 1000, fontface = 3) +
      geom_label_repel(data = cap1_2,  #Adlercreutzia equolifaciens"
            stat = "identity",
            mapping = aes(x = PCoA1/3, y = PCoA2/3, label=cap1_2$shortaxa),
            segment.size  = 0.5,
            segment.color = "#A9A9A9",
            direction = "both",
            size = 4, nudge_x = 0.2, nudge_y = -0.02, 
            inherit.aes = F, 
            point.padding = 0.5,
            seed = 987, max.iter = 1000, fontface = 3) +
  geom_label_repel(data = cap1_19, 
            stat = "identity",
            mapping = aes(x = PCoA1/3, y = PCoA2/3, label=cap1_19$shortaxa),
            segment.size  = 0.5,
            segment.color = "#A9A9A9",
            direction = "both",
            size = 4, nudge_x = -0.15, nudge_y = -0.04, 
            inherit.aes = F, 
            point.padding = 0.5,
            seed = 987, max.iter = 1000, fontface = 3) +
    geom_label_repel(data = cap1_3,  #"Akkermansia muciniphila"
            stat = "identity",
            mapping = aes(x = PCoA1/3, y = PCoA2/3, label=cap1_3$shortaxa),
            segment.size  = 0.5,
            segment.color = "#A9A9A9",
            direction = "both",
            size = 4, nudge_x = 0.05, nudge_y = 0.1, 
            inherit.aes = F, 
            point.padding = 0.5,
            seed = 987, max.iter = 1000, fontface = 3) +
      geom_label_repel(data = cap1_4,  #"Asaccharobacter celatus"
            stat = "identity",
            mapping = aes(x = PCoA1/3, y = PCoA2/3, label=cap1_4$shortaxa),
            segment.size  = 0.5,
            segment.color = "#A9A9A9",
            direction = "both",
            size = 4, nudge_x = 0.5, nudge_y = 0, 
            inherit.aes = F, 
            point.padding = 0.5,
            seed = 987, max.iter = 1000, fontface = 3) +
      geom_label_repel(data = cap1_7,  #"Bacteroides caecimuris"
            stat = "identity",
            mapping = aes(x = PCoA1/3, y = PCoA2/3, label=cap1_7$shortaxa),
            segment.size  = 0.5,
            segment.color = "#A9A9A9",
            direction = "both",
            size = 4, nudge_x = 0.01, nudge_y = 0.1, 
            inherit.aes = F, 
            point.padding = 0.5,
            seed = 987, max.iter = 1000, fontface = 3) +
      geom_label_repel(data = cap1_6,  #"Enterorhabdus caecimuris
            stat = "identity",
            mapping = aes(x = PCoA1/3, y = PCoA2/3, label=cap1_6$shortaxa),
            segment.size  = 0.5,
            segment.color = "#A9A9A9",
            direction = "both",
            size = 4, nudge_x = 0.5, nudge_y = -0.01, 
            inherit.aes = F, 
            point.padding = 0.5,
            seed = 987, max.iter = 1000, fontface = 3) +
    geom_label_repel(data = cap1_5,  #"Muribaculaceae sp DSM 103720"
            stat = "identity",
            mapping = aes(x = PCoA1/3, y = PCoA2/3, label=cap1_5$shortaxa),
            segment.size  = 0.5,
            segment.color = "#A9A9A9",
            direction = "both",
            size = 4, nudge_x = 0.1, nudge_y = -00.1, 
            inherit.aes = F, 
            point.padding = 0.5,
            seed = 987, max.iter = 1000, fontface = 3) +
        geom_label_repel(data = cap1_8,  #"Bacteroides uniformis"
            stat = "identity",
            mapping = aes(x = PCoA1/3, y = PCoA2/3, label=cap1_8$shortaxa),
            segment.size  = 0.5,
            segment.color = "#A9A9A9",
            direction = "both",
            size = 4, nudge_x = 0.2, nudge_y = 0.03, 
            inherit.aes = F, 
            point.padding = 0.5,
            seed = 987, max.iter = 1000, fontface = 3) +
          geom_label_repel(data = cap1_9,  #"Parabacteroides distasonis"
            stat = "identity",
            mapping = aes(x = PCoA1/3, y = PCoA2/3, label=cap1_9$shortaxa),
            segment.size  = 0.5,
            segment.color = "#A9A9A9",
            direction = "both",
            size = 4, nudge_x = 0, nudge_y = 0.2, 
            inherit.aes = F, 
            point.padding = 0.5,
            seed = 987, max.iter = 1000, fontface = 3) +
        geom_label_repel(data = cap1_10,  #"Enterococcus faecalis"
            stat = "identity",
            mapping = aes(x = PCoA1/3, y = PCoA2/3, label=cap1_10$shortaxa),
            segment.size  = 0.5,
            segment.color = "#A9A9A9",
            direction = "both",
            size = 4, nudge_x = -0.05, nudge_y = -0.02, 
            inherit.aes = F, 
            point.padding = 0.5,
            seed = 987, max.iter = 1000, fontface = 3) +
      geom_label_repel(data = cap1_11,  #"Lactobacillus johnsonii"
            stat = "identity",
            mapping = aes(x = PCoA1/3, y = PCoA2/3, label=cap1_11$shortaxa),
            segment.size  = 0.5,
            segment.color = "#A9A9A9",
            direction = "both",
            size = 4, nudge_x = -0.1, nudge_y = 0.05, 
            inherit.aes = F, 
            point.padding = 0.5,
            seed = 987, max.iter = 1000, fontface = 3) +
        geom_label_repel(data = cap1_12,  #"Acutalibacter muris"
            stat = "identity",
            mapping = aes(x = PCoA1/3, y = PCoA2/3, label=cap1_12$shortaxa),
            segment.size  = 0.5,
            segment.color = "#A9A9A9",
            direction = "both",
            size = 4, nudge_x = -0.1, nudge_y = -0.5, 
            inherit.aes = F, 
            point.padding = 0.5,
            seed = 987, max.iter = 1000, fontface = 3) +
        geom_label_repel(data = cap1_13,  #"Clostridium cocleatum"
            stat = "identity",
            mapping = aes(x = PCoA1/3, y = PCoA2/3, label=cap1_13$shortaxa),
            segment.size  = 0.5,
            segment.color = "#A9A9A9",
            direction = "both",
            size = 4, nudge_x = -0.02, nudge_y = 0.05, 
            inherit.aes = F, 
            point.padding = 0.5,
            seed = 987, max.iter = 1000, fontface = 3) +
        geom_label_repel(data = cap1_14,  #"Prevotella sp MGM2"
            stat = "identity",
            mapping = aes(x = PCoA1/3, y = PCoA2/3, label=cap1_14$shortaxa),
            segment.size  = 0.5,
            segment.color = "#A9A9A9",
            direction = "both",
            size = 4, nudge_x = -0.1, nudge_y = 0.02, 
            inherit.aes = F, 
            point.padding = 0.5,
            seed = 987, max.iter = 1000, fontface = 3) +
      geom_label_repel(data = cap1_15,  #"Faecalibaculum rodentium"
            stat = "identity",
            mapping = aes(x = PCoA1/3, y = PCoA2/3, label=cap1_15$shortaxa),
            segment.size  = 0.5,
            segment.color = "#A9A9A9",
            direction = "both",
            size = 4, nudge_x = 0, nudge_y = 0.2, 
            inherit.aes = F, 
            point.padding = 0.5,
            seed = 987, max.iter = 1000, fontface = 3) +
       geom_label_repel(data = cap1_16,  #"Parabacteroides goldsteinii"
            stat = "identity",
            mapping = aes(x = PCoA1/3, y = PCoA2/3, label=cap1_16$shortaxa),
            segment.size  = 0.5,
            segment.color = "#A9A9A9",
            direction = "both",
            size = 4, nudge_x = -0.05, nudge_y = 0.05, 
            inherit.aes = F, 
            point.padding = 0.5,
            seed = 987, max.iter = 1000, fontface = 3) +
      geom_label_repel(data = cap1_17,  #"Oscillibacter sp 1 3"
            stat = "identity",
            mapping = aes(x = PCoA1/3, y = PCoA2/3, label=cap1_17$shortaxa),
            segment.size  = 0.5,
            segment.color = "#A9A9A9",
            direction = "both",
            size = 4, nudge_x = -0.1, nudge_y = -0.02, 
            inherit.aes = F, 
            point.padding = 0.5,
            seed = 987, max.iter = 1000, fontface = 3) +
        geom_label_repel(data = cap1_18,  #"Escherichia coli"
            stat = "identity",
            mapping = aes(x = PCoA1/3, y = PCoA2/3, label=cap1_18$shortaxa),
            segment.size  = 0.5,
            segment.color = "#A9A9A9",
            direction = "both",
            size = 4, nudge_x = -0.05, nudge_y = 0.01, 
            inherit.aes = F, 
            point.padding = 0.5,
            seed = 987, max.iter = 1000, fontface = 3) +
       geom_label_repel(data = cap1_20,  #Dorea
            stat = "identity",
            mapping = aes(x = PCoA1/3, y = PCoA2/3, label=cap1_20$shortaxa),
            segment.size  = 0.5,
            segment.color = "#A9A9A9",
            direction = "both",
            size = 4, nudge_x = 0.05, nudge_y = 0.01, 
            inherit.aes = F, 
            point.padding = 0.5,
            seed = 987, max.iter = 1000, fontface = 3) 
gss_prepost<-gs +   annotate("label", label = "PRE-FMT YO", x = -0.12, y = -0.16, size = 4.5, colour = "black",fontface = 2) +
  annotate("label", label = "PRE-FMT OY", x = -0.15, y = -0.025, size = 4.5, colour = "black",fontface = 2) +
  annotate("label", label = "PRE-FMT YY", x = -0.03, y = -0.085, size = 4.5, colour = "black",fontface = 2) +
  annotate("label", label = "POST-FMT YO", x = -0.05, y = 0.1, size = 4.5, colour = "black",fontface = 2) +
  annotate("label", label = "POST-FMT YY", x = 0.275, y = -0.08, size = 4.5, colour = "black",fontface = 2) +
  annotate("label", label = "POST-FMT OY", x = 0.25, y = 0.1, size = 4.5, colour = "black",fontface = 2) 
gss_prepost

ggsave(gss_prepost, filename = "Taxonomy/Figures/PCoA_pre_post_FMT_pluspecies.svg", 
       device = "svg", width = 8, height = 8)
```


# PCoA for start plus Procrustes analyses

*START* timepoint _Pre-Abx_
```{r, message=FALSE, warning=FALSE}
pre_abx_phylo<-subset_samples(sp_phylo, Timepoint=="PRE_Abx")
pre_abx_phylo<-filter_taxa(pre_abx_phylo, function(x) sum(x) > 0, TRUE)


# calculate JSD
sample_names(pre_abx_phylo)<-gsub(".*_","", sample_names(pre_abx_phylo))
pre_abx_jsd<-phyloseq::distance(pre_abx_phylo, method= "jsd", na.rm=F)

# Create df
all(rownames(t(otu_table(pre_abx_phylo))) == rownames(meta(pre_abx_phylo)))
df_m_preabx<-cbind.data.frame(t(otu_table(pre_abx_phylo)), meta(pre_abx_phylo))

# PCoA
bd_preabx<-betadisper(pre_abx_jsd, group = df_m_preabx$receiver, add = T, type = "centroid")
plot(bd_preabx)
```

Make a PCoA plot including species
```{r, message=FALSE, warning=FALSE}
ord_meta<-bd_preabx$vectors[,1:2]
all(rownames(df_m_preabx) == rownames(ord_meta))
ord_meta<-cbind.data.frame(ord_meta, df_m_preabx)


# Calculate proportion of variacne explained 
eig<-bd_preabx$eig
axis1<-paste("Axis 1 [", round(eig[1]/sum(eig),2)*100, "%]", sep = "")
axis2<-paste("Axis 2 [", round(eig[2]/sum(eig),2)*100, "%]", sep = "")

# Order the factors of age as Y, M and O
ord_meta$receiver<-factor(ord_meta$receiver, levels = c("Y", "M", "O"))

# Make a graph
(g_preabx<-ggplot(ord_meta, aes(PCoA1, PCoA2, col = receiver)) +
  geom_convexhull(alpha = 0.3,aes(fill = receiver, color = receiver), show.legend = F) +
  scale_color_manual(values = rep("black", 9)) +
  scale_fill_brewer(palette = "RdYlBu") +
  new_scale_color() +  
  geom_point(color = "black",size = 4.5) +
  geom_point(aes(color = receiver),size = 3.5) +
  scale_color_brewer(palette = "RdYlBu") +
  labs(title = "Gut microbiota structure at the beginning of the experiment across mice ages", shape = "Receiver", color = "Donor") +
  xlab(axis1) +
  ylab(axis2) +
  theme_bw() +
  theme(plot.title = element_text(size = 13, face = "bold"), legend.position = "none", panel.grid = element_blank()))

sf <- envfit(bd_preabx, as.data.frame(t(otu_table(pre_abx_phylo))), 
             display = "sites", permutations = 10000) 

# Get the scores for plotting
vec.sp.df.scores<-as.data.frame(scores(sf, display = "vector"))
ef.df<-as.data.frame(sf$vectors$arrows*sqrt(sf$vectors$r))
ef.df$species<-rownames(ef.df)

# Collecting only taxa with significant pvalues
A <- as.list(sf$vectors)
pvals<-as.data.frame(A$pvals)
arrows<-as.data.frame(A$arrows*sqrt(A$r))
C<-cbind(arrows, pvals)

C$padj<-p.adjust(C$`A$pvals`, method = "fdr")
Cred<-subset(C,padj<0.05)
Cred <- cbind(Cred, Family = rownames(Cred))

# Collecting only Species with the highest degree of variation along the axis
cap1<-Cred
taxa<-rownames(cap1)
shortaxa<-taxa
for(i in 1:length(shortaxa)){
  x<-strsplit(shortaxa[i], split = "|", fixed = T)[[1]][7]
  x<-gsub("s__", "", x)
  shortaxa[i]<-gsub("_", " ", x)

}

cap1$shortaxa<-shortaxa
library(ggrepel)
gs_preabx<- g_preabx +   geom_segment(data = cap1, stat = "identity",
              mapping = aes(x = 0, xend = PCoA1/3, y = 0, yend = PCoA2/3),
              arrow = arrow(length = unit(2, "mm"), type = "open"), 
              colour = "#C0C0C0", 
              inherit.aes = F, 
              linejoin = "mitre",
              arrow.fill = "#C0C0C0") +
  geom_label_repel(data = cap1,  
            stat = "identity",
            mapping = aes(x = PCoA1/3, y = PCoA2/3, label=cap1$shortaxa),
            segment.size  = 0.5,
            segment.color = "#A9A9A9",
            direction = "both",
            size = 4, nudge_x = 0, nudge_y = 0.00, 
            inherit.aes = F, 
            point.padding = 0.5,
            seed = 987, max.iter = 1000, fontface = 3, force = 10)

gss_preabx<-gs_preabx + annotate("label", label = "Y", x = -0.05, y = -0.2, size = 8, colour = "black",fontface = 2) +
  annotate("label", label = "M", x = -0.2, y = 0.15, size = 8, colour = "black",fontface = 2) +
  annotate("label", label = "O", x = -0.3, y = -0.05, size = 8, colour = "black",fontface = 2) 
gss_preabx
ggsave(gss_preabx, filename = "Taxonomy/Figures/PCoA_preAbx.svg", 
       device = "svg", width = 8, height = 8)
```

*Test significance of clustering*
```{r}
set.seed(5698)
# The adjustment is already included
(pair_pre_abx<-pairwise.adonis(pre_abx_jsd, df_m_preabx$receiver, 
                                       p.adjust.m = "BH", perm = 10000))
```

*After antiobiotic treat* 
timepoint _POST-ABX_
```{r}
post_abx_phylo<-subset_samples(sp_phylo, Timepoint=="POST_ABX")
post_abx_phylo<-filter_taxa(post_abx_phylo, function(x) sum(x) > 0, TRUE)

# calculate JSD
sample_names(post_abx_phylo)<-gsub(".*_","", sample_names(post_abx_phylo))
post_abx_jsd<-phyloseq::distance(post_abx_phylo, method= "jsd", na.rm=F)

# Create df
all(rownames(t(otu_table(post_abx_phylo))) == rownames(meta(post_abx_phylo)))
df_m_postabx<-cbind.data.frame(t(otu_table(post_abx_phylo)), meta(post_abx_phylo))

# PCoA
bd_postabx<-betadisper(post_abx_jsd, group = df_m_postabx$receiver, add = T, type = "centroid")
plot(bd_postabx)
```


*After FMT* 
timepoint _POST-FMT_
```{r, message=FALSE, warning=FALSE}
post_fmt_phylo<-subset_samples(sp_phylo, Timepoint=="POST_FMT")
post_fmt_phylo<-filter_taxa(post_fmt_phylo, function(x) sum(x) > 0, TRUE)

# calculate JSD
sample_names(post_fmt_phylo)<-gsub(".*_","", sample_names(post_fmt_phylo))
post_fmt_jsd<-phyloseq::distance(post_fmt_phylo, method= "jsd", na.rm=F)

# Create df 
all(rownames(t(otu_table(post_fmt_phylo))) == rownames(meta(post_fmt_phylo)))
df_m_postfmt<-cbind.data.frame(t(otu_table(post_fmt_phylo)), meta(post_fmt_phylo))

# PCoA
bd_postfmt<-betadisper(post_fmt_jsd, group = df_m_postfmt$groups, add = T, type = "centroid")
plot(bd_postfmt)
```


*End of the experiment* 
timepoint _END_
```{r, message=FALSE, warning=FALSE}
end_phylo<-subset_samples(sp_phylo, Timepoint=="END")
end_phylo<-filter_taxa(end_phylo, function(x) sum(x) > 0, TRUE)

# calculate JSD
sample_names(end_phylo)<-gsub(".*_","", sample_names(end_phylo))
end_jsd<-phyloseq::distance(end_phylo, method= "jsd", na.rm=F)

# Create df
all(rownames(t(otu_table(end_phylo))) == rownames(meta(end_phylo)))
df_m_end<-cbind.data.frame(t(otu_table(end_phylo)), meta(end_phylo))

# PCoA
bd_end<-betadisper(end_jsd, group = df_m_end$groups, add = T, type = "centroid")
plot(bd_end)
```

## PROTEST

Only matching samples need to be selected.

*START*

START vs POST_ABX
```{R}
sample_start<-bd_preabx$distances %>% names
sample_postabx<-bd_postabx$distances %>% names

sort(sample_start)
sort(sample_postabx)

match_s<-sample_start[sample_start %in% sample_postabx]
no_match<-sample_start[!(sample_start %in% sample_postabx)]

all(match_s %in% sample_start)
all(match_s %in% sample_postabx)

match_phylo<-subset_samples(sp_phylo, mice %in% match_s)

start<-subset_samples(match_phylo, Timepoint=="PRE_Abx")
start_jsd<-phyloseq::distance(start, method= "jsd", na.rm=F)
start_bd<-ordinate(start, method = "PCoA", distance = start_jsd)

po<-subset_samples(match_phylo, Timepoint=="POST_ABX")
po_jsd<-phyloseq::distance(po, method= "jsd", na.rm=F)
po_bd<-ordinate(po, method = "PCoA", distance = po_jsd)

set.seed(56987)
protest( po_bd$vectors[,1:2], start_bd$vectors[,1:2], permutations = 10000)

```

START vs POST_FMT
```{R}
sample_postfmt<-bd_postfmt$distances %>% names

sort(sample_start)
sort(sample_postfmt)

match_s<-sample_start[sample_start %in% sample_postabx]
no_match<-sample_start[!(sample_start %in% sample_postabx)]

all(match_s %in% sample_start)
# These are the shared samples.

match_phylo<-subset_samples(sp_phylo, mice %in% match_s)

pof<-subset_samples(match_phylo, Timepoint=="POST_FMT")
pof_jsd<-phyloseq::distance(pof, method= "jsd", na.rm=F)
pof_bd<-ordinate(pof, method = "PCoA", distance = pof_jsd)


protest(pof_bd$vectors[,1:2], start_bd$vectors[,1:2],  permutations = 10000)

```

START vs END
```{R}
sample_end<-bd_end$distances %>% names

sort(sample_start)
sort(sample_end)

match_s<-sample_end[sample_end %in% sample_start]
match_s2<-sample_start[sample_start %in% sample_end]
match_s<-unique(match_s, match_s2)
no_match_missing_start<-sample_end[!(sample_end %in% sample_start)]
no_match_missing_end<-sample_start[!(sample_start %in% sample_end)]

all(match_s %in% sample_start)
all(match_s %in% sample_postabx)

match_phylo<-subset_samples(sp_phylo, mice %in% match_s)

start<-subset_samples(match_phylo, Timepoint=="PRE_Abx")
start_jsd<-phyloseq::distance(start, method= "jsd", na.rm=F)
start_bd<-ordinate(start, method = "PCoA", distance = start_jsd)

end_e<-subset_samples(match_phylo, Timepoint=="END")
end_jsd<-phyloseq::distance(end_e, method= "jsd", na.rm=F)
end_bd<-ordinate(end_e, method = "PCoA", distance = end_jsd)


set.seed(56987)
protest(end_bd$vectors[,1:2], start_bd$vectors[,1:2], permutations = 10000)
```


*POST-Abx*

POST_ABX vs POST_FMT
```{R}

sort(sample_postabx)
sort(sample_postfmt)
all(sort(sample_postabx) == sort(sample_postfmt))

po_bd<-ordinate(post_abx_phylo, method = "PCoA", distance = post_abx_jsd)
pof_bd<-ordinate(post_fmt_phylo, method = "PCoA", distance = post_fmt_jsd)


set.seed(56987)
protest(po_bd$vectors[,1:2], pof_bd$vectors[,1:2], permutations = 10000)

```

POST_ABX vs END
```{R}

sort(sample_postabx)
sort(sample_end)
all(sort(sample_postabx) == sort(sample_end))



match_s<-sample_end[sample_end %in% sample_postabx]
match_s2<-sample_postabx[sample_postabx %in% sample_end]
match_s<-unique(match_s, match_s2)
no_match_missing_start<-sample_end[!(sample_end %in% sample_postabx)]
no_match_missing_end<-sample_postabx[!(sample_postabx %in% sample_end)]


all(match_s %in% sample_end)
all(match_s %in% sample_postabx)

match_phylo<-subset_samples(sp_phylo, mice %in% match_s)

end<-subset_samples(match_phylo, Timepoint=="END")
end_jsd<-phyloseq::distance(end, method= "jsd", na.rm=F)
end_bd<-ordinate(end, method = "PCoA", distance = end_jsd)

po<-subset_samples(match_phylo, Timepoint=="POST_ABX")
po_jsd<-phyloseq::distance(po, method= "jsd", na.rm=F)
po_bd<-ordinate(po, method = "PCoA", distance = po_jsd)


set.seed(56987)
protest(end_bd$vectors[,1:2], po_bd$vectors[,1:2], ermutations = 10000)

```

*POST-FMT*

POST_FMT vs END
```{R}

sort(sample_postfmt)
sort(sample_end)
all(sort(sample_postfmt) == sort(sample_end))



match_s<-sample_end[sample_end %in% sample_postfmt]
match_s2<-sample_postfmt[sample_postfmt %in% sample_end]
match_s<-unique(match_s, match_s2)
no_match_missing_start<-sample_end[!(sample_end %in% sample_postfmt)]
no_match_missing_end<-sample_postfmt[!(sample_postfmt %in% sample_end)]


all(match_s %in% sample_end)
all(match_s %in% sample_postfmt)

match_phylo<-subset_samples(sp_phylo, mice %in% match_s)

end<-subset_samples(match_phylo, Timepoint=="END")
end_jsd<-phyloseq::distance(end, method= "jsd", na.rm=F)
end_bd<-ordinate(end, method = "PCoA", distance = end_jsd)

pof<-subset_samples(match_phylo, Timepoint=="POST_FMT")
pof_jsd<-phyloseq::distance(pof, method= "jsd", na.rm=F)
pof_bd<-ordinate(pof, method = "PCoA", distance = pof_jsd)


set.seed(56987)
protest(end_bd$vectors[,1:2], pof_bd$vectors[,1:2], ermutations = 100000)

```

## Save dataset

```{r}
dir.create("Taxonomy/RData")
save.image("Taxonomy/RData/00_Load_clean_beta.RData")
```

