---
title: "DA at start"
author: "Stefano Romano"
date: "October 2020"
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

# Load the needed packages

Load all required packages and data.


```{r, message=F}
list.of.packages <- c("ggplot2", "dplyr", "reshape2", "ggnewscale",
                      "propr", "devtools", "ggpubr", "vegan", "RColorBrewer", "car", 
                      "DHARMa", "glmmTMB", "lme4", "papeR")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

if (!requireNamespace("BiocManager", quietly = TRUE)){
  install.packages("BiocManager")}


list.of.bioc <- c("phyloseq", "grDevices", "microbiome")
new.packages <- list.of.bioc[!(list.of.bioc %in% installed.packages()[,"Package"])]
if(length(new.packages)) BiocManager::install(new.packages)

library(RColorBrewer)
library(car)
library(lme4)
library(phyloseq)
library(microbiome)
library(ggpubr)
library(ggnewscale)
library(ggplot2)
library(reshape2)
library(grDevices)
library(vegan)
library(devtools)
library(CoDaSeq)
library(ggrepel)
library(glmmTMB)
library(DHARMa)
library(dplyr)

phyl_yy_in<-readRDS("Taxonomy/RDS/Phyloseq_filtered_species.rds")
```

Select only samples for the correct timepoint: _PRE-ABX_
```{r}
phyl_yy_in<-subset_samples(phyl_yy_in, Timepoint == "PRE_Abx")
```

Verify whether there are taxa with re. abund of 0 on all samples

```{r}
ntaxa(phyl_yy_in)
x<-genefilter_sample(phyl_yy_in, filterfun_sample(function(x) x == 0), A=1*nsamples(phyl_yy_in))
x<-!(x)
phyl_yy<-prune_taxa(x, phyl_yy_in)
ntaxa(phyl_yy)
```

Keep only taxe that have a rel. abund. > 0 in at least 10% of the samples.
We check the data manully and we verified that this treshold was the most appropriate for the data.

```{R}
# Keep only species that ahave rel abund > 0 in at least 10% samples
wh0 <- genefilter_sample(phyl_yy, filterfun_sample(function(x) x > 0), A=0.1*nsamples(phyl_yy))
phyl_yy <- prune_taxa(wh0, phyl_yy)
ntaxa(phyl_yy)
x<-taxa_names(phyl_yy_in)
setdiff(x,taxa_names(phyl_yy))
```

Create a new df including rel. ab. and metadata

```{r}
all(rownames(as.data.frame(t(otu_table(phyl_yy)))) == rownames(meta(phyl_yy)))
# data can be merged
df_m<-cbind.data.frame(as.data.frame(t(otu_table(phyl_yy))), meta(phyl_yy))
tax_name<-taxa_names(phyl_yy)
```

Look at the minimum values to correct the transformation
```{R}
otu_table(phyl_yy) %>% as.vector() %>% sort(decreasing = F) %>% unique() %>% head
```


Total number of transformation tested

- LOG 
- LOG add 0 after (LOG zero)
- CLR
- ArcSin 
- LOGIT 


# Run full models and look how they are 
```{R  message=FALSE, warning=FALSE}
# Create lists of lmm for each taxa
l_glm_logbefore<-vector(mode = "list", length = length(tax_name))
names(l_glm_logbefore)<-tax_name
l_glm_logZero<-l_glm_logbefore
l_glm_clr<-l_glm_logbefore
l_glm_arcsinbefore<-l_glm_logbefore
l_glm_logitbefore<-l_glm_logbefore

# make clr
df_clr<-df_m[,tax_name]
df_clr[df_clr == 0]<-1e-06
df_clr<-as.data.frame(codaSeq.clr(df_clr))
print(all(rownames(df_m) == rownames(df_clr)))
df_clr<-cbind.data.frame(df_clr, df_m)

for(i in 1:length(tax_name)){
  # LOG 
  l_before<-log2(df_m[,tax_name[i]]+1e-06) 
  # LOG ZERO
  l_zero<-log2(df_m[,tax_name[i]])
  l_zero[is.na(l_zero) | is.infinite(l_zero)]<-0
  # ArcSin
  arcs_before<-asin(sqrt(df_m[,tax_name[i]]+1e-06))
  # LOGIT
  logit_before<-logit(df_m[,tax_name[i]]+1e-06) 
  # Create the models
  print(tax_name[i])
  print("Log")
  l_glm_logbefore[[i]]<-lmer(l_before ~ receiver + (1|cage_mice), data = df_m)
  print("Log zero")
  l_glm_logZero[[i]]<-lmer(l_zero ~ receiver + (1|cage_mice), data = df_m)
  print("CLR")
  l_glm_clr[[i]]<-lmer(df_clr[,i] ~ receiver + (1|cage_mice), data = df_clr)
  print("ArcSin")
  l_glm_arcsinbefore[[i]]<-lmer(arcs_before ~ receiver + (1|cage_mice), data = df_m)
  print("LOGIT")
  l_glm_logitbefore[[i]]<-lmer(logit_before ~ receiver + (1|cage_mice), data = df_m)
 
}
```

Save residuals.

```{r message=FALSE, warning=FALSE, paged.print=FALSE}

for(i in 1:length(tax_name)){
  # log 
  path<-"Taxonomy/Figures/Residuals_species_across_age_only_preAbx_cagerandom/LOG/"
  dir.create("Taxonomy/Figures/Residuals_species_across_age_only_preAbx_cagerandom/LOG", recursive = T)
  x<-paste0(path, tax_name[i], ".png", sep = "")
  trellis.device(device="png", file = x)
  print(plot(l_glm_logbefore[[i]]))
  dev.off()
  png(paste0(path, tax_name[i], "_qq.png", sep = ""))
  qqnorm(residuals(l_glm_logbefore[[i]]))
  qqline(residuals(l_glm_logbefore[[i]]))
  dev.off()
  # LOG ZERO
  path<-"Taxonomy/Figures/Residuals_species_across_age_only_preAbx_cagerandom/LOGZERO/"
  dir.create("Taxonomy/Figures/Residuals_species_across_age_only_preAbx_cagerandom/LOGZERO", recursive = T)
  x<-paste0(path, tax_name[i], ".png", sep = "")
  trellis.device(device="png", file = x)
  print(plot(l_glm_logZero[[i]]))
  dev.off()
  png(paste0(path, tax_name[i], "_qq.png", sep = ""))
  qqnorm(residuals(l_glm_logZero[[i]]))
  qqline(residuals(l_glm_logZero[[i]]))
  dev.off()  
  # CLR
  dir.create("Taxonomy/Figures/Residuals_species_across_age_only_preAbx_cagerandom/CLR/", recursive = T)
  path<-"Taxonomy/Figures/Residuals_species_across_age_only_preAbx_cagerandom/CLR/"
  x<-paste0(path, tax_name[i], ".png", sep = "")
  trellis.device(device="png", file = x)
  print(plot(l_glm_clr[[i]]))
  dev.off()
  png(paste0(path, tax_name[i], "_qq.png", sep = ""))
  qqnorm(residuals(l_glm_clr[[i]]))
  qqline(residuals(l_glm_clr[[i]]))
  dev.off()  
  # ArcSin 
  dir.create("Taxonomy/Figures/Residuals_species_across_age_only_preAbx_cagerandom/ArcSin/", recursive = T)
  path<-"Taxonomy/Figures/Residuals_species_across_age_only_preAbx_cagerandom/ArcSin/"
  x<-paste0(path, tax_name[i], ".png", sep = "")
  trellis.device(device="png", file = x)
  print(plot(l_glm_arcsinbefore[[i]]))
  dev.off()
  png(paste0(path, tax_name[i], "_qq.png", sep = ""))
  qqnorm(residuals(l_glm_arcsinbefore[[i]]))
  qqline(residuals(l_glm_arcsinbefore[[i]]))
  dev.off()  
  # LOGIT 
  dir.create("Taxonomy/Figures/Residuals_species_across_age_only_preAbx_cagerandom/LOGIT/", recursive = T)
  path<-"Taxonomy/Figures/Residuals_species_across_age_only_preAbx_cagerandom/LOGIT/"
  x<-paste0(path, tax_name[i], ".png", sep = "")
  trellis.device(device="png", file = x)
  print(plot(l_glm_logitbefore[[i]]))
  dev.off()
  png(paste0(path, tax_name[i], "_qq.png", sep = ""))
  qqnorm(residuals(l_glm_logitbefore[[i]]))
  qqline(residuals(l_glm_logitbefore[[i]]))
  dev.off() 
}
```

Residuals are checked manually.
The transformations that result in the best residuals are selected and saved below.

## Extract only the contrast of interest
We will verify whether the specific contrasts are significant.
We will use the lmm with the best residuals for this.
For all significant taxa, we will then collect the CLR values and report them in a graph.

```{R message=FALSE, warning=FALSE}
library(emmeans)
emm_options(contrast = list(infer = c(TRUE, T)))

# Set up contrasts
y<-c(1,0,0)
o<-c(0,1,0)

# CLR
l_contr<-l_glm_clr
for(i in 1:length(l_contr)){
  print(i)
  emm1<-emmeans(l_glm_clr[[i]], specs = ~ receiver)
  l_contr[[i]]<-contrast(emm1, method = list("Y - O" = y - o), adjust = "none") %>% as.data.frame()}

df_contr <- do.call(rbind.data.frame, l_contr)
df_contr$padjust<-p.adjust(df_contr$p.value, method = "fdr")

# ArcSin
# Subset only relevant taxa
relevant_arc<-c("k__Bacteria|p__Actinobacteria|c__Coriobacteriia|o__Eggerthellales|f__Eggerthellaceae|g__Adlercreutzia|s__Adlercreutzia_equolifaciens",
            "k__Bacteria|p__Actinobacteria|c__Coriobacteriia|o__Eggerthellales|f__Eggerthellaceae|g__Asaccharobacter|s__Asaccharobacter_celatus",
            "k__Bacteria|p__Bacteroidetes|c__Bacteroidia|o__Bacteroidales|f__Bacteroidaceae|g__Bacteroides|s__Bacteroides_caecimuris",
            "k__Bacteria|p__Firmicutes|c__Bacilli|o__Lactobacillales|f__Enterococcaceae|g__Enterococcus|s__Enterococcus_faecalis",
            "k__Bacteria|p__Firmicutes|c__Bacilli|o__Lactobacillales|f__Lactobacillaceae|g__Lactobacillus|s__Lactobacillus_reuteri",
            "k__Bacteria|p__Firmicutes|c__Clostridia|o__Clostridiales|f__Lachnospiraceae|g__Lachnospiraceae_unclassified|s__Lachnospiraceae_bacterium_3_2",
            "k__Bacteria|p__Proteobacteria|c__Gammaproteobacteria|o__Enterobacterales|f__Enterobacteriaceae|g__Escherichia|s__Escherichia_coli")                                                            

l_contr_arc<-l_glm_arcsinbefore
for(i in 1:length(l_contr_arc)){
  print(i)
  emm1<-emmeans(l_glm_arcsinbefore[[i]], specs = ~ receiver)
  l_contr_arc[[i]]<-contrast(emm1, method = list("Y - O" = y - o), adjust = "none") %>% as.data.frame()}

df_contr_arc <- do.call(rbind.data.frame, l_contr_arc)
df_contr_arc$padjust<-p.adjust(df_contr_arc$p.value, method = "fdr")
all(relevant_arc %in% rownames(df_contr_arc))

df_contr_arc<-df_contr_arc[which(rownames(df_contr_arc) %in% relevant_arc),]
arc_sign<-rownames(subset(df_contr_arc,  padjust <= 0.05))

# LOGZERO
# Subset only relevant taxa
relevant_logzero<-c("k__Bacteria|p__Bacteroidetes|c__Bacteroidia|o__Bacteroidales|f__Bacteroidaceae|g__Bacteroides|s__Bacteroides_thetaiotaomicron",
            "k__Bacteria|p__Deferribacteres|c__Deferribacteres|o__Deferribacterales|f__Deferribacteraceae|g__Mucispirillum|s__Mucispirillum_schaedleri",
            "k__Bacteria|p__Firmicutes|c__Bacilli|o__Lactobacillales|f__Lactobacillaceae|g__Lactobacillus|s__Lactobacillus_murinus",
            "k__Bacteria|p__Firmicutes|c__Clostridia|o__Clostridiales|f__Lachnospiraceae|g__Acetatifactor|s__Acetatifactor_muris",
            "k__Bacteria|p__Firmicutes|c__Clostridia|o__Clostridiales|f__Ruminococcaceae|g__Acutalibacter|s__Acutalibacter_muris",
            "k__Bacteria|p__Firmicutes|c__Clostridia|o__Clostridiales|f__Lachnospiraceae|g__Lachnospiraceae_unclassified|s__Lachnospiraceae_bacterium_28_4",
            "k__Bacteria|p__Firmicutes|c__Erysipelotrichia|o__Erysipelotrichales|f__Erysipelotrichaceae|g__Faecalibaculum|s__Faecalibaculum_rodentium")                                                


l_contr_log<-l_glm_logZero
for(i in 1:length(l_contr_log)){
  print(i)
  emm1<-emmeans(l_glm_logZero[[i]], specs = ~ receiver)
  l_contr_log[[i]]<-contrast(emm1, method = list("Y - O" = y - o), adjust = "none") %>% as.data.frame()}

df_contr_log <- do.call(rbind.data.frame, l_contr_log)
df_contr_log$padjust<-p.adjust(df_contr_log$p.value, method = "fdr")
all(relevant_logzero %in% rownames(df_contr_log))

df_contr_log<-df_contr_log[which(rownames(df_contr_log) %in% relevant_logzero),]
log_sign<-rownames(subset(df_contr_log,  padjust <= 0.05))


```


Select only the taxa w p-value <= 0.05

```{R message=FALSE, warning=FALSE}
all(c(relevant_logzero, relevant_arc) %in% rownames(df_contr))

# Remove from df_clr the taxa relevant in the other transformations
nrow(df_contr)
df_contr_sign<-df_contr[-1*(which(rownames(df_contr) %in% c(relevant_logzero, relevant_arc))),]
nrow(df_contr_sign)
c(relevant_logzero, relevant_arc) %>% length()

# Select the significant in df clr
df_contr_sign<-subset(df_contr_sign, padjust <= 0.05)

# Check that all species in other transf are in the df clr
c(log_sign, arc_sign)
# Merge sign taxa
df_contr_sign<-rbind.data.frame(df_contr_sign, df_contr[which(rownames(df_contr) %in% c(log_sign, arc_sign)),])

shortaxa<-rownames(df_contr_sign)

for(i in 1:length(shortaxa)){
  x<-strsplit(shortaxa[i], split = "|", fixed = T)[[1]][7]
  x<-gsub("s__", "", x)
  shortaxa[i]<-gsub("_", " ", x)
}
shortaxa<-gsub("[.][0-9]*", "", shortaxa)
df_contr_sign$short<-shortaxa

df_contr_sign$short<-as.factor(df_contr_sign$short)
for(i in 1:nrow(df_contr_sign)){
  if(df_contr_sign[i,]$estimate < 0){
    df_contr_sign$enrich[i]<-"Enriched O"
  } else {df_contr_sign$enrich[i]<-"Enriched Y"}
}
df_contr_sign$enrich <- factor(df_contr_sign$enrich, levels = c("Enriched O", "Enriched Y"))


gg_clr<-ggplot(df_contr_sign, aes(reorder(short, -estimate), estimate*-1, fill = enrich)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  facet_wrap(~contrast) +
  geom_hline(yintercept=0, linetype="dashed") +
  scale_fill_grey() +
  geom_errorbar(aes(ymin = (estimate - SE)*-1, ymax =   (estimate + SE) *-1), width = 0.2) +
  ylab("Mean difference in CLR and SE") +
  xlab("Species") +
  theme_bw() +
  theme(legend.position = "bottom", axis.text = element_text(size = 14), panel.background = element_blank(), panel.grid = element_blank(), axis.title = element_text(size =18)) 

gg_clr

ggsave("Taxonomy/Figures/CLR_onlyfdr_DA_Pre_Abx.svg", device = "svg", width = 8, height = 8)

```

Create a table including the info about the normalization method used.
```{R message=FALSE, warning=FALSE}
df_contr_sign<-df_contr[-1*(which(rownames(df_contr) %in% c(relevant_logzero, relevant_arc))),]
df_contr_sign<-subset(df_contr_sign, padjust <= 0.05)
df_contr_sign$Transformation<-rep("CLR", length = nrow(df_contr_sign))

df_log_sign<-subset(df_contr_log,  padjust <= 0.05)
df_log_sign$Transformation<-rep("LogZero", length = nrow(df_log_sign))


df_arcsin_sign<-subset(df_contr_arc,  padjust <= 0.05)
df_arcsin_sign$Transformation<-rep("ArcSin", length = nrow(df_arcsin_sign))
df_transf_sign<-rbind.data.frame(df_contr_sign, df_log_sign, df_arcsin_sign)
  
  
dir.create("Taxonomy/Tables/", recursive = T)
write.table(df_transf_sign, "Taxonomy/Tables/DA_species_Pre_Abx.txt", sep = ",")
write.table(df_contr, "Taxonomy/Tables/DA_allSpecies_Pre_Abx.txt", sep = ",")

```

### Save image
```{r}
save.image("Taxonomy/RData/07_DA_lm_onlyAbx_cagerandom.RData")
```
