---
title: "Load and filter functional data"
author: "Stefano Romano"
date: "October, 2020"
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

# Load the needed packages

Load all required packages and the previous workspace.

```{r, message=F}
list.of.packages <- c("ggplot2", "dplyr", "reshape2",
                      "propr", "devtools", "ggpubr", "vegan", "RColorBrewer", "car", 
                      "DHARMa", "glmmTMB", "lme4", "papeR")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

if (!requireNamespace("BiocManager", quietly = TRUE)){
  install.packages("BiocManager")}


list.of.bioc <- c("phyloseq", "grDevices", "microbiome")
new.packages <- list.of.bioc[!(list.of.bioc %in% installed.packages()[,"Package"])]
if(length(new.packages)) BiocManager::install(new.packages)

library(RColorBrewer)
library(car)
library(lme4)
library(ggConvexHull)
library(phyloseq)
library(microbiome)
library(ggpubr)
library(ggplot2)
library(reshape2)
library(grDevices)
library(vegan)
library(devtools)
library(CoDaSeq)
library(ggrepel)
library(pairwiseAdonis)
library(ggnewscale)
library(glmmTMB)
library(DHARMa)
library(dplyr)
```

# Clean dataset

Load the pathway information.
Select only the relevant samples.

```{R}
# Make phyloseq object
phyl_yy<-phyloseq(otu_table(path, taxa_are_rows = F), sample_data(m))
ntaxa(phyl_yy)
```


# Clean the dataset

```{R}
x<-genefilter_sample(phyl_yy, filterfun_sample(function(x) x == 0), A=1*nsamples(phyl_yy))
x<-!(x)
phyl_yy<-prune_taxa(x, phyl_yy)
ntaxa(phyl_yy)
```


### Do graphs with single pathway

```{R echo=TRUE}
# Create formatted df
all(rownames(as.data.frame(otu_table(phyl_yy))) == rownames(meta(phyl_yy)))
df_m<-cbind.data.frame(as.data.frame(otu_table(phyl_yy)), meta(phyl_yy))

# Select proper paths names
t<-taxa_names(phyl_yy)
all(t == names(df_m)[1:396])

shortaxa<-gsub(".*: ", "" ,t)
shortaxa<-gsub("/", "-" ,shortaxa)
df_m$Timepoint<-factor(df_m$Timepoint, levels = c("PRE_Abx", "POST_ABX", "POST_FMT", "END"))
pa<-"/Functions/Figures/Time_resolved_path_abund/"
dir.create(pa)
for(i in 1:length(t)){
  ggplot(df_m, aes(Timepoint, log2(df_m[,i]), color = Timepoint)) +
    geom_boxplot(outlier.colour = NULL) +
    scale_colour_brewer(type = "qual", palette = 3)  +
    new_scale_color() +
    geom_jitter(width = 0.2, aes(fill = cage), pch = 21, alpha = 0.8, size = 2.5) +
    scale_fill_brewer(type = "qual", palette = 1)  +
    theme_bw() +
    xlab("Time points") +
    ylab("log2(rel. abund)")+
    labs(title = paste0("Change in relative abundance of the path ", shortaxa[i])) +
    theme(axis.text.x = element_text(angle = 90)) +
    facet_wrap(~groups)
  ggsave(paste0(pa, shortaxa[i], ".png"), device = "png", width = 10)
  
}
```

Keep all paths present in all samples of OY or YO at Start or Post_FMT

#### Check for Old
```{r}
### Pre Abx
phylo_oy<-subset_samples(phyl_yy, Timepoint == "PRE_Abx" & groups == "O Y ")
# All mice
wh0_pre <- genefilter_sample(phylo_oy, filterfun_sample(function(x) x  > 0), A=1*nsamples(phylo_oy))
sum(wh0_pre)

## POST FMT
phylo_oy<-subset_samples(phyl_yy, Timepoint == "POST_FMT" & groups == "O Y ")
# All mice
wh0_post <- genefilter_sample(phylo_oy, filterfun_sample(function(x) x  > 0), A=1*nsamples(phylo_oy))
sum(wh0_post)
oy_to_keep<-c(wh0_pre[wh0_pre == T], wh0_post[wh0_post == T])
```

#### Check for Young
```{r}
### Pre Abx
phylo_oy<-subset_samples(phyl_yy, Timepoint == "PRE_Abx" & groups == "Y O ")
# All mice
wh0_pre <- genefilter_sample(phylo_oy, filterfun_sample(function(x) x  > 0), A=1*nsamples(phylo_oy))
sum(wh0_pre)

## POST FMT
phylo_oy<-subset_samples(phyl_yy, Timepoint == "POST_FMT" & groups == "Y O ")
# All mice
wh0_post <- genefilter_sample(phylo_oy, filterfun_sample(function(x) x  > 0), A=1*nsamples(phylo_oy))
sum(wh0_post)

yo_to_keep<-c(wh0_pre[wh0_pre == T], wh0_post[wh0_post == T])
```

*Merge missing and filtered paths*
```{r}
# Combine the paths to keep in a single vector
to_keep_oy_yo<-unique(names(c(yo_to_keep, oy_to_keep)))


to_add<-as.logical(rep(TRUE, length(to_keep_oy_yo)))
names(to_add)<-to_keep_oy_yo

# select all paths in the filtered dataset
to_remove_names<-setdiff(taxa_names(phyl_yy), to_keep_oy_yo)
to_remove<-as.logical(rep(FALSE, length(to_remove_names)))
names(to_remove)<-to_remove_names


tota_genes_to_keep_pluremove<-c(to_remove,to_add)
names(tota_genes_to_keep_pluremove) <-as.character(names(tota_genes_to_keep_pluremove))

all(sort(names(tota_genes_to_keep_pluremove)) == sort(taxa_names(phyl_yy)))
```

There is a problem with 5' in names
```{r}
all(to_keep_oy_yo %in% colnames(path))
names(path)<-gsub("5'", "5", names(path))
all(to_keep_oy_yo %in% colnames(path))

path_2<-path[,which(colnames(path) %in% to_keep_oy_yo)]
ncol(path_2)

# all filtered path are now selected in the df
phyl_yy_filtered<-phyloseq(otu_table(path_2, taxa_are_rows = F), sample_data(m))
```


Save data

```{R}
saveRDS(phyl_yy_filtered, "Functions/RDS/Phyloseq_filtered_path_yooy.rds")
```

## Filter only path in all at Start/End YO OY

#### Chek for Y
```{r}
### Pre Abx
phylo_oy<-subset_samples(phyl_yy, Timepoint == "PRE_Abx" & receiver == "Y")
# All mice
wh0_pre <- genefilter_sample(phylo_oy, filterfun_sample(function(x) x  > 0), A=0.9*nsamples(phylo_oy))
sum(wh0_pre)

## POST FMT
phylo_oy<-subset_samples(phyl_yy, Timepoint == "PRE_Abx" & receiver == "O")
# All mice
wh0_post <- genefilter_sample(phylo_oy, filterfun_sample(function(x) x  > 0), A=0.9*nsamples(phylo_oy))
sum(wh0_post)

oy_to_keep<-c(wh0_pre[wh0_pre == T], wh0_post[wh0_post == T])
```

Merge missing and filtered paths.

```{r}
to_keep_oy_yo<-unique(names(c(yo_to_keep, oy_to_keep)))

to_add<-as.logical(rep(TRUE, length(to_keep_oy_yo)))
names(to_add)<-to_keep_oy_yo
# select all paths in the filtered dataset
to_remove_names<-taxa_names(phyl_yy)[which(!(taxa_names(phyl_yy) %in% to_keep_oy_yo))]
to_remove<-rep(FALSE, length(to_remove_names))
names(to_remove)<-to_remove_names
tota_genes_to_keep_pluremove<-c(to_remove,to_add)
```

```{r}
all(to_keep_oy_yo %in% names(path))
names(path)<-gsub("5'", "5", names(path))
all(to_keep_oy_yo %in% names(path))

path_2<-path[,which(colnames(path) %in% to_keep_oy_yo)]
phyl_yy_filtered_start<-phyloseq(otu_table(path_2, taxa_are_rows = F), sample_data(m))
```


### Save data and image
```{R}
saveRDS(phyl_yy_filtered_start, "Functions/RDS/Phyloseq_filtered_path_start.rds")
save.image("Functions/RData/02_Filter_format_DA.RData")

```

