---
title: "DA lmm, pathways"
author: "Stefano Romano"
date: "October 2020"
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

# Load the needed packages

Load all required packages and data.

```{r, message=F}
list.of.packages <- c("ggplot2", "dplyr", "reshape2", "reshape2",
                      "propr", "devtools", "ggpubr", "vegan", "RColorBrewer", "car", 
                      "DHARMa", "glmmTMB", "lme4", "papeR")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

if (!requireNamespace("BiocManager", quietly = TRUE)){
  install.packages("BiocManager")}


list.of.bioc <- c("phyloseq", "grDevices", "microbiome")
new.packages <- list.of.bioc[!(list.of.bioc %in% installed.packages()[,"Package"])]
if(length(new.packages)) BiocManager::install(new.packages)

library(RColorBrewer)
library(car)
library(lme4)
library(phyloseq)
library(microbiome)
library(ggpubr)
library(ggplot2)
library(reshape2)
library(grDevices)
library(vegan)
library(devtools)
library(CoDaSeq)
library(ggrepel)
library(ggnewscale)
library(glmmTMB)
library(DHARMa)
library(dplyr)
library(emmeans)

phyl_yy<-readRDS("Functions//RDS/Phyloseq_filtered_path_yooy.rds")
ntaxa(phyl_yy)
```

Data will be transformed using different approaches and linear mixed model will be built.
We will manually assess the quality of the models.

```{r}
all(rownames(as.data.frame(otu_table(phyl_yy))) == rownames(meta(phyl_yy)))
df_m<-cbind.data.frame(as.data.frame(otu_table(phyl_yy)), meta(phyl_yy))
tax_name<-taxa_names(phyl_yy)
```

Look at the minimum values to correct data before transformation.
```{R}
otu_table(phyl_yy) %>% as.vector() %>% sort(decreasing = F) %>% unique() %>% head
```

Total number of transformation tested

- LOG 
- LOG add 0 after (LOG zero)
- CLR
- ArcSin 
- LOGIT 
- No transform


# Run full models 
```{R}
# Create lists of lmm for each path
l_glm_logbefore<-vector(mode = "list", length = length(tax_name))
names(l_glm_logbefore)<-tax_name
l_glm_logZero<-l_glm_logbefore
l_glm_clr<-l_glm_logbefore
l_glm_arcsinbefore<-l_glm_logbefore
l_glm_logitbefore<-l_glm_logbefore
l_glm_notransf<-l_glm_logbefore

# make clr
df_clr<-df_m[,tax_name]
df_clr[df_clr == 0]<-1e-09
df_clr<-as.data.frame(codaSeq.clr(df_clr))
print(all(rownames(df_m) == rownames(df_clr)))
df_clr<-cbind.data.frame(df_clr, df_m)

all(tax_name == names(df_m)[1:280])
all(tax_name == names(df_clr)[1:280])


for(i in 1:length(tax_name)){
  # LOG 
  l_before<-log2(df_m[,tax_name[i]]+1e-09) 
  # LOG ZERO
  l_zero<-log2(df_m[,tax_name[i]])
  l_zero[is.na(l_zero) | is.infinite(l_zero)]<-0
  # ArcSin 
  arcs_before<-asin(sqrt(df_m[,tax_name[i]]+1e-09))
  # LOGIT 
  logit_before<-logit(df_m[,tax_name[i]]+1e-09) 
  # Create the models
  print(tax_name[i])
  print("Log")
  l_glm_logbefore[[i]]<-lmer(l_before ~ receiver * treat * Timepoint + (1|cage_mice/mice), data = df_m)
   print("Log zero")
  l_glm_logZero[[i]]<-lmer(l_zero ~ receiver * treat * Timepoint +  (1|cage_mice/ mice), data = df_m)
  print("CLR")
  l_glm_clr[[i]]<-lmer(df_clr[,i] ~ receiver * treat * Timepoint +  (1|cage_mice/mice), data = df_clr)
  print("ArcSin")
  l_glm_arcsinbefore[[i]]<-lmer(arcs_before ~ receiver * treat * Timepoint +  (1|cage_mice/mice), data = df_m)
  print("LOGIT")
  l_glm_logitbefore[[i]]<-lmer(logit_before ~ receiver * treat * Timepoint +  (1|cage_mice/mice), data = df_m)
  print("No Transf")
  l_glm_notransf[[i]]<-lmer(df_m[,i] ~ receiver * treat * Timepoint +  (1|cage_mice/mice), data = df_m)

}
```

Save residuals.
```{r message=FALSE, warning=FALSE, paged.print=TRUE}

for(i in 1:length(tax_name)){
  # log
  path<-"/Functions//Figures/Residuals_path_full_model/LOG/"
  dir.create("Functions/Figures/Residuals_path_full_model/LOG", recursive = T)
  x<-paste0(path, gsub("/", "", tax_name[i]), ".png", sep = "")
  trellis.device(device="png", file = x)
  print(plot(l_glm_logbefore[[i]]))
  dev.off()
  png(paste0(path, gsub("/", "",tax_name[i]), "_qq.png", sep = ""))
  qqnorm(residuals(l_glm_logbefore[[i]]))
  qqline(residuals(l_glm_logbefore[[i]]))
  dev.off()
  # LOG ZERO
  path<-"Functions/Figures/Residuals_path_full_model/LOGZERO/"
  dir.create("Functions/Figures/Residuals_path_full_model/LOGZERO", recursive = T)
  x<-paste0(path, gsub("/", "", tax_name[i]), ".png", sep = "")
  trellis.device(device="png", file = x)
  print(plot(l_glm_logZero[[i]]))
  dev.off()
  png(paste0(path, gsub("/", "",tax_name[i]), "_qq.png", sep = ""))
  qqnorm(residuals(l_glm_logZero[[i]]))
  qqline(residuals(l_glm_logZero[[i]]))
  dev.off()  
  # CLR
  dir.create("Functions/Figures/Residuals_path_full_model/CLR/", recursive = T)
  path<-"Functions/Figures/Residuals_path_full_model/CLR/"
  x<-paste0(path, gsub("/", "", tax_name[i]), ".png", sep = "")
  trellis.device(device="png", file = x)
  print(plot(l_glm_clr[[i]]))
  dev.off()
  png(paste0(path, gsub("/", "",tax_name[i]), "_qq.png", sep = ""))
  qqnorm(residuals(l_glm_clr[[i]]))
  qqline(residuals(l_glm_clr[[i]]))
  dev.off()  
  # ArcSin 
  dir.create("Functions/Figures/Residuals_path_full_model/ArcSin/", recursive = T)
  path<-"Functions/Figures/Residuals_path_full_model/ArcSin/"
  x<-paste0(path, gsub("/", "", tax_name[i]), ".png", sep = "")
  trellis.device(device="png", file = x)
  print(plot(l_glm_arcsinbefore[[i]]))
  dev.off()
  png(paste0(path, gsub("/", "",tax_name[i]), "_qq.png", sep = ""))
  qqnorm(residuals(l_glm_arcsinbefore[[i]]))
  qqline(residuals(l_glm_arcsinbefore[[i]]))
  dev.off()  
  # LOGIT
  dir.create("Functions/Figures/Residuals_path_full_model/LOGIT/", recursive = T)
  path<-"Functions/Figures/Residuals_path_full_model/LOGIT/"
  x<-paste0(path, gsub("/", "", tax_name[i]), ".png", sep = "")
  trellis.device(device="png", file = x)
  print(plot(l_glm_logitbefore[[i]]))
  dev.off()
  png(paste0(path, gsub("/", "",tax_name[i]), "_qq.png", sep = ""))
  qqnorm(residuals(l_glm_logitbefore[[i]]))
  qqline(residuals(l_glm_logitbefore[[i]]))
  dev.off() 
  # No Trans
  dir.create("Functions/Figures/Residuals_path_full_model/NoTransf/", recursive = T)
  path<-"Functions/Figures/Residuals_path_full_model/NoTransf/"
  x<-paste0(path, gsub("/", "", tax_name[i]), ".png", sep = "")
  trellis.device(device="png", file = x)
  print(plot(l_glm_notransf[[i]]))
  dev.off()
  png(paste0(path, gsub("/", "",tax_name[i]), "_qq.png", sep = ""))
  qqnorm(residuals(l_glm_notransf[[i]]))
  qqline(residuals(l_glm_notransf[[i]]))
  dev.off() 
}
```

Residuals are checked manually.
The transformations that result in the best residuals are selected and saved in a table.
This table is then imported and formatted as below.

```{r}
res<-read.delim("Functions/Tables/Residuals.txt", header = T)
# Check how many model types are good
tax_name_shor<-gsub(": .*", "", tax_name)
res$short<-gsub(": .*", "", res$Path)
all(tax_name_shor %in% res$short)

res<-subset(res, short %in% tax_name_shor)
nrow(res)
all(sort(res$short) == sort(tax_name_shor))

# Select only the ones for the diff transf
res$Best_model<-as.character(res$Best_model)

res_clr<-gsub(": .*", "", res[res$Best_model=="CLR",]$Path)
res_not<-gsub(": .*", "", res[res$Best_model=="NoTran",]$Path)
res_log_zero<-gsub(": .*", "", res[res$Best_model=="LogZero",]$Path)
res_arcsin<-gsub(": .*", "", res[res$Best_model=="ArcSin",]$Path)
res_logit<-gsub(": .*", "", res[res$Best_model=="LogitBefore",]$Path)

sum(length(res_clr), length(res_arcsin), length(res_not), length(res_log_zero), length(res_logit))

```


## Extract only the contrast of interest
We will verify whether the specific contrasts are significant.
We will use the lmm with the best residuals for this.
For all significant paths, we will then collect the CLR values and report them in a graph.

```{R }
emm_options(contrast = list(infer = c(TRUE, T)))

# create custom contrasts
s_yy<-c(1,rep(0,15))
pfmt_yy<-c(rep(0,8),1,rep(0,7))
s_ym<-c(rep(0,4),1,rep(0,11))
pfmt_ym<-c(rep(0,12),1,rep(0,3))
s_oy<-c(0,1,rep(0,14))
pfmt_oy<-c(rep(0,9),1,rep(0,6))
s_yo<-c(rep(0,2),1,rep(0,13))
pfmt_yo<-c(rep(0,10),1,rep(0,5))

# NON transformed
if( c("UNINTEGRATED") %in% names(l_glm_notransf)){
  l_glm_notransf<-l_glm_notransf[-1*which(names(l_glm_notransf) %in% c("UNINTEGRATED","UNMAPPED"))]
}

l_contr_notransf<-l_glm_notransf
for(i in 1:length(l_glm_notransf)){
  print(i)
  emm1<-emmeans(l_glm_notransf[[i]], specs = ~ receiver:treat:Timepoint)
  l_contr_notransf[[i]]<-contrast(emm1, method = list("S_OY-PFMT_OY" = s_oy - pfmt_oy,
                             "S_YO-PFMT_YO" = s_yo - pfmt_yo), adjust = "none") %>% as.data.frame()
}
df_contr_nontranf <- do.call(rbind.data.frame, l_contr_notransf)
df_contr_nontranf$padjust<-p.adjust(df_contr_nontranf$p.value, method = "fdr")
df_contr_nontranf$short<-gsub(": .*", "", gsub("[.][0-9]*", "", rownames(df_contr_nontranf)))
all(res_not %in% df_contr_nontranf$short)

df_contr_nontranf<-subset(df_contr_nontranf, short %in% res_not)
length(res_not) == nrow(df_contr_nontranf)/2

notr_sign<-rownames(subset(df_contr_nontranf, padjust <= 0.05))

# ArcSin
if( c("UNINTEGRATED") %in% names(l_glm_arcsinbefore)){
  l_glm_arcsinbefore<-l_glm_arcsinbefore[-1*which(names(l_glm_arcsinbefore) %in% c("UNINTEGRATED","UNMAPPED"))]
}

l_contr_arcsinbefore<-l_glm_arcsinbefore
for(i in 1:length(l_contr_arcsinbefore)){
  print(i)
  emm1<-emmeans(l_glm_arcsinbefore[[i]], specs = ~ receiver:treat:Timepoint)
  l_contr_arcsinbefore[[i]]<-contrast(regrid(emm1), method = list("S_OY-PFMT_OY" = s_oy - pfmt_oy,
                             "S_YO-PFMT_YO" = s_yo - pfmt_yo), adjust = "none") %>% as.data.frame()
}
df_contr_arcsinbefore <- do.call(rbind.data.frame, l_contr_arcsinbefore)
df_contr_arcsinbefore$padjust<-p.adjust(df_contr_arcsinbefore$p.value, method = "fdr")
df_contr_arcsinbefore$short<-gsub(": .*", "", gsub("[.][0-9]*", "", rownames(df_contr_arcsinbefore)))
all(res_arcsin %in% df_contr_arcsinbefore$short)

df_contr_arcsinbefore<-subset(df_contr_arcsinbefore, short %in% res_arcsin)
length(res_arcsin) == nrow(df_contr_arcsinbefore)/2

arc_sign<-rownames(subset(df_contr_arcsinbefore, padjust <= 0.05))

# LogZero
if( c("UNINTEGRATED") %in% names(l_glm_logZero)){
  l_glm_logZero<-l_glm_logZero[-1*which(names(l_glm_logZero) %in% c("UNINTEGRATED","UNMAPPED"))]
}

l_contr_logZero<-l_glm_logZero
for(i in 1:length(l_contr_logZero)){
  print(i)
  emm1<-emmeans(l_glm_logZero[[i]], specs = ~ receiver:treat:Timepoint)
  l_contr_logZero[[i]]<-contrast(regrid(emm1), method = list("S_OY-PFMT_OY" = s_oy - pfmt_oy,
                             "S_YO-PFMT_YO" = s_yo - pfmt_yo), adjust = "none") %>% as.data.frame()
}
df_contr_logZero <- do.call(rbind.data.frame, l_contr_logZero)
df_contr_logZero$padjust<-p.adjust(df_contr_logZero$p.value, method = "fdr")
df_contr_logZero$short<-gsub(": .*", "", gsub("[.][0-9]*", "", rownames(df_contr_logZero)))
all(res_log_zero %in% df_contr_logZero$short)

df_contr_logZero<-subset(df_contr_logZero, short %in% res_log_zero)
length(res_log_zero) == nrow(df_contr_logZero)/2

logZero_sign<-rownames(subset(df_contr_logZero, padjust <= 0.05))

# Logit
if( c("UNINTEGRATED") %in% names(l_glm_logitbefore)){
  l_glm_logitbefore<-l_glm_logitbefore[-1*which(names(l_glm_logitbefore) %in% c("UNINTEGRATED","UNMAPPED"))]
}

l_contr_logitbefore<-l_glm_logitbefore
for(i in 1:length(l_contr_logitbefore)){
  print(i)
  emm1<-emmeans(l_glm_logitbefore[[i]], specs = ~ receiver:treat:Timepoint)
  l_contr_logitbefore[[i]]<-contrast(regrid(emm1), method = list("S_OY-PFMT_OY" = s_oy - pfmt_oy,
                             "S_YO-PFMT_YO" = s_yo - pfmt_yo), adjust = "none") %>% as.data.frame()
}
df_contr_logitbefore <- do.call(rbind.data.frame, l_contr_logitbefore)
df_contr_logitbefore$padjust<-p.adjust(df_contr_logitbefore$p.value, method = "fdr")
df_contr_logitbefore$short<-gsub(": .*", "", gsub("[.][0-9]*", "", rownames(df_contr_logitbefore)))
all(res_logit %in% df_contr_logitbefore$short)

df_contr_logitbefore<-subset(df_contr_logitbefore, short %in% res_logit)
length(res_logit) == nrow(df_contr_logitbefore)/2

logitbefore_sign<-rownames(subset(df_contr_logitbefore, padjust <= 0.05))


# CLR
if( c("UNINTEGRATED") %in% names(l_glm_clr)){
  l_glm_clr<-l_glm_clr[-1*which(names(l_glm_clr) %in% c("UNINTEGRATED","UNMAPPED"))]
}

l_contr<-l_glm_clr
for(i in 1:length(l_contr)){
  print(i)
  emm1<-emmeans(l_glm_clr[[i]], specs = ~ receiver:treat:Timepoint)
  l_contr[[i]]<-contrast(emm1, method = list("S_OY-PFMT_OY" = s_oy - pfmt_oy,
                             "S_YO-PFMT_YO" = s_yo - pfmt_yo), adjust = "none") %>% as.data.frame()
}
df_contr <- do.call(rbind.data.frame, l_contr)
df_contr$padjust<-p.adjust(df_contr$p.value, method = "fdr")
df_contr$short<-gsub(": .*", "", gsub("[.][0-9]*", "", rownames(df_contr)))
all(res_clr %in% df_contr$short)

df_contr_clr<-subset(df_contr, short %in% res_clr)
length(res_clr) == nrow(df_contr_clr)/2

df_contr_sign<-subset(df_contr_clr,  padjust <= 0.05)

# Check all sign are present in clr table
other_tr_sign<-c(notr_sign, arc_sign, logZero_sign, logitbefore_sign)
all( c(notr_sign, arc_sign, logZero_sign, logitbefore_sign) %in% rownames(df_contr))

df_contr_sign<-rbind.data.frame(df_contr_sign, df_contr[which(rownames(df_contr) %in% c(notr_sign, arc_sign, logZero_sign, logitbefore_sign)),])
nrow(df_contr_sign)

```

Select only sign taxa.

```{R}
df_contr_sign$short<-as.factor(df_contr_sign$short)
df_contr_sign$path<-gsub("[.][0-9]", "", gsub(".*: ", "", rownames(df_contr_sign)))
df_contr_sign$path<-as.factor(df_contr_sign$path)

for(i in 1:nrow(df_contr_sign)){
  if(df_contr_sign[i,]$estimate < 0){
    df_contr_sign$enrich[i]<-"Enriched in POST_FMT"
  } else {df_contr_sign$enrich[i]<-"Enriched in PRE_FMT"}
}
df_contr_sign$enrich <- factor(df_contr_sign$enrich, levels = c("Enriched in PRE_FMT", "Enriched in POST_FMT"))

gg_clr<-ggplot(df_contr_sign, aes(reorder(path, -estimate), estimate*-1, fill = enrich)) +
  geom_bar(stat = "identity") +
  coord_flip() +
  facet_wrap(~contrast) +
  geom_hline(yintercept=0, linetype="dashed") +
  scale_fill_brewer(type = "div", palette = "Set2") +
  geom_errorbar(aes(ymin = -1*(estimate-SE), ymax =-1*(estimate + SE)), width = 0.2) +
  labs(title = "Pathways showing a significant difference in abundance across conditions") +
  theme_bw() +
  theme(legend.position = "bottom") +
  ylab("Mean difference in CLR and SE") +
  xlab("path")

ggsave("Functions/Figures/CLR_onlyfdr_DA.svg", device = "svg", height = 100, width = 25, units = "cm")
```

Save a table with the info about the different transf.
```{r}

no_tr_df<-subset(df_contr_nontranf, padjust <= 0.05)
no_tr_df$Transformation<-rep("Non transformed", length = nrow(no_tr_df))

arcsin_df<-subset(df_contr_arcsinbefore, padjust <= 0.05)
arcsin_df$Transformation<-rep("ArcSin", length = nrow(arcsin_df))

logZero_df<-subset(df_contr_logZero, padjust <= 0.05)
logZero_df$Transformation<-rep("LogZero", length = nrow(logZero_df))
                     
logitbefore_sign<-subset(df_contr_logitbefore, padjust <= 0.05)
logitbefore_sign$Transformation<-rep("Logit", length = nrow(logitbefore_sign))
                           
df_contr_sign<-subset(df_contr_clr,  padjust <= 0.05)
df_contr_sign$Transformation<-rep("CLR", length = nrow(df_contr_sign))


df_transf_sing<-rbind.data.frame(no_tr_df, arcsin_df, logZero_df, logitbefore_sign, df_contr_sign)
dir.create("/Functions/Tables/", recursive = T)
write.table(df_transf_sing, "Functions/Tables/DA_pathways_Sign.txt", sep = ",")
write.table(df_contr, "Functions/Tables/DA_allPathways.txt", sep = ",")

```

### Save image
```{r}
save.image("/Functions/RData/03_DA_lm.RData")
```
