---
title: "Metabolome ordination and PLS-DA"
author: "Stefano Romano"
date: "January 2021"
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

# Load the needed packages

Load all required packages and the previous workspace.


```{r, message=F}
list.of.packages <- c("ggplot2", "dplyr", "reshape2",
                      "ggpubr", "vegan", "RColorBrewer", "car")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

if (!requireNamespace("BiocManager", quietly = TRUE)){
  install.packages("BiocManager")}


list.of.bioc <- c("phyloseq", "grDevices", "microbiome", "mixOmics")
new.packages <- list.of.bioc[!(list.of.bioc %in% installed.packages()[,"Package"])]
if(length(new.packages)) BiocManager::install(new.packages)

library(RColorBrewer)
library(mixOmics)
library(car)
library(ggConvexHull)
library(phyloseq)
library(microbiome)
library(ggpubr)
library(ggplot2)
library(reshape2)
library(grDevices)
library(vegan)
library(ggrepel)
# If pairwiseAdonis is not installed, please install the package before loading the library.
library(pairwiseAdonis)
library(dplyr)
```

# Clean and reformat the dataset

Load the metabolome. 

```{R}

met<-read.delim("metabolome.txt", 
                header = F, fill = T)
```
Transform the data.
```{r}
as.matrix(met) %>% as.vector() %>% sort() %>% unique() %>% head()
# Change factors to numbers
str(met)
met[]<-lapply(met, function(x) as.numeric(as.character(x)))
str(met)

met[met == 0] <- 0.001
met_log<-log2(met)
head(met_log)
```

Center and divide by standard deviation.

```{r}
met_log<-as.data.frame(t(met_log))
met_log_scaled<-scale(as.matrix(met_log), center = T, scale = T)
met_log_scaled_df<-as.data.frame(met_log_scaled)

# Visually verify distributions
df_m_met<-melt(met_log_scaled_df)
ggplot(df_m_met, aes(variable, value)) +
  geom_boxplot() +
  coord_flip()
```


Format df.
```{r}
met_log_scaled_df$samples<-gsub("\ [1-9]*", "", rownames(met_log_scaled_df))
# add cage info
all(rownames(met_log_scaled_df) == cage$codes)
met_log_scaled_df$cage<-cage$cage

# Make values as factors
met_log_scaled_df$samples<-as.factor(met_log_scaled_df$samples)
met_log_scaled_df$cage<-as.factor(met_log_scaled_df$cage)
```

Make explorative plot.
```{r}
vegdist(met_log_scaled_df[, -1*(c(which(names(met_log_scaled_df) == "samples"), which(names(met_log_scaled_df) == "cage")))], method = "euclidean") -> euc_met
bd<-betadisper(euc_met, group = met_log_scaled_df$samples, add = T, type = "centroid")
plot(bd)
```

There is one sample ("Pre Abx_OY 611") that does not cluster together with any other sample.
This is most likely an outlier and we will remove it from the dataset.

```{r}
df<-met_log_scaled_df[rownames(met_log_scaled_df) != "Pre Abx_OY 611",]
df$samples<-factor(df$samples, levels = c("PreAbx_YO", "PreAbx_OY", "PostFMT_OY", "PostFMT_YO"))
```

## PCoA

```{r}
vegdist(df[, -1*(c(which(names(df) == "samples"), which(names(df) == "cage")))], method = "euclidean") -> euc_met
bd<-betadisper(euc_met, group = df$samples, add = T, type = "centroid")
plot(bd)
```

We will not consider the outlier further in the analyses.

# Verify whether differences between samples are significant

```{r}
set.seed(987)
cc<-capscale(euc_met ~ samples + Condition(cage), data = df)

h <- how(blocks = df$cage, nperm = 10000)
anova.cca(cc, data = df, permutations = h, parallel = 2)

# Pairwise adonis
l_p<-pairwise.adonis2(euc_met ~ samples, data = df, strata = 'cage')
p_values<-lapply(l_p[-1], function(x) x$`Pr(>F)`[1])
p.adjust(p_values, method = "BH")
```

There is a statistically significant difference between groups.
This difference is just partially mantained (adjusted p-values < 0.1) when we perform a pairwise analysis. 

```{r}
summary(cc)
```


# PLS-DA

PLS-DA will be performed using the mixOmics package.

## Run model including tuning

Run a model with 10 components.

```{r}
set.seed(2543)
Y<-as.factor(df$samples)
X<-as.matrix(df[,1:67])

MyResult.splsda <-splsda(X, Y, ncomp = 10, near.zero.var=TRUE) 
perf.plsda <- perf(MyResult.splsda, validation = "Mfold", folds = 5, 
                  progressBar = F, auc = TRUE, nrepeat = 75, cpus = 2) 

plot(perf.plsda, col = color.mixo(1:3), sd = TRUE, legend.position = "horizontal")

```

The classification error rate is not particularly low.
With 3 components the error rate goes down to 30%.

Use this 3 components and perfom a variable selection.

```{r}
# grid of possible keepX values that will be tested for each comp 
list.keepX <- c(1:5, seq(10, 67, 5))
set.seed(2543)
tune.splsda <- tune.splsda(X, Y, ncomp = 3, 
                           validation = 'Mfold', 
                           folds = 5, 
                           progressBar = F, 
                           dist = 'max.dist',
                           test.keepX = list.keepX, 
                           nrepeat = 75, 
                           measure = "BER",
                           cpus = 2)

error <- tune.splsda$error.rate  # error rate per component for the keepX grid
error
ncomp <- tune.splsda$choice.ncomp$ncomp # optimal number of components based on t-tests
ncomp

select.keepX <- tune.splsda$choice.keepX[1:ncomp]  # optimal number of variables to select
select.keepX

plot(tune.splsda, col = color.jet(3))
```


## Make the final plot

```{r}

MyResult.splsda <-splsda(X, Y, ncomp = ncomp, keepX = select.keepX,
                         near.zero.var=TRUE) 
# Asses performance
set.seed(40)
perf.srbct <- perf(MyResult.splsda, validation = "Mfold", folds = 5,
                   dist = 'max.dist', nrepeat = 75,
                   progressBar = F, cpus = 2) 
perf.srbct$error.rate
plot(perf.srbct, col = color.mixo(5))

### Plot graphs
pls12<-plotIndiv(MyResult.splsda, ind.names = FALSE, legend=TRUE, comp = c(1,2),
          ellipse = TRUE, star = TRUE, title = 'sPLS-DA comp 1 & 2',
          X.label = 'PLS-DA 1', Y.label = 'PLS-DA 2')
pls13<-plotIndiv(MyResult.splsda, ind.names = FALSE, legend=TRUE, comp = c(1,3),
          ellipse = TRUE, star = TRUE, title = 'sPLS-DA comp 1 & 3',
          X.label = 'PLS-DA 1', Y.label = 'PLS-DA 3')

gg<-ggarrange(pls12[[3]], pls13[[3]], common.legend = T)
ggsave(gg, filename = "Metabolome/Figures/PLS_DA_1_2.svg", device = "svg",  
       width = 14, height = 7.5)

# Loadings
# Need to be saved manually
plotLoadings(MyResult.splsda, comp = 1, title = 'Loadings on component 1', 
             contrib = 'max', method = 'mean', size.name = 1, name.var.complete = T,size.title = 1.5, border = T)
plotLoadings(MyResult.splsda, comp = 2, title = 'Loadings on component 2', 
             contrib = 'max', method = 'mean', size.name = 0.75, name.var.complete = T,size.title = 1.5, border = T)
plotLoadings(MyResult.splsda, comp = 3, title = 'Loadings on component 3', 
             contrib = 'max', method = 'mean', size.name = 0.75, name.var.complete = T,size.title = 1.5, border = T)
```

Figure have then been manually formatted to make them consistent with rest of the paper.

Save the workspace.
```{r}
save.image("Metabolome/RData/Metabolome.RData")
```