---
title: "DA metabolome"
author: "Stefano Romano"
date: "October 2020"
output:
  html_document: default
  pdf_document: default
editor_options:
  chunk_output_type: console
---

# Load the needed packages

Load all required packages.

```{r, message=F}
list.of.packages <- c("ggplot2", "dplyr", "reshape2",
                      "propr", "devtools", "ggpubr", "vegan", "RColorBrewer", "car", 
                      "DHARMa", "glmmTMB", "lme4", "papeR")
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
if(length(new.packages)) install.packages(new.packages)

if (!requireNamespace("BiocManager", quietly = TRUE)){
  install.packages("BiocManager")}


list.of.bioc <- c("phyloseq", "grDevices", "microbiome", "mixOmics")
new.packages <- list.of.bioc[!(list.of.bioc %in% installed.packages()[,"Package"])]
if(length(new.packages)) BiocManager::install(new.packages)

library(RColorBrewer)
library(mixOmics)
library(car)
library(lme4)
library(ggConvexHull)
library(phyloseq)
library(microbiome)
library(ggpubr)
library(ggplot2)
library(reshape2)
library(grDevices)
library(vegan)
library(devtools)
library(CoDaSeq)
library(ggrepel)
library(ggnewscale)
library(glmmTMB)
library(DHARMa)
library(dplyr)
```

# Load and pre-pocess data

Load the metabolome.
```{R}

met<-read.delim("Raw_data/metabolome.txt", 
                header = F, fill = T)
```

Transform the data.

```{r}
# Identify lowest constant to add to 0
as.matrix(met) %>% as.vector() %>% sort() %>% unique() %>% head()

# Convert values in numeric
str(met)
met[]<-lapply(met, function(x) as.numeric(as.character(x)))
str(met)

met[met == 0] <- 0.001
met_log<-log2(met)
head(met_log)

```

Center and divide by standard deviation.

```{r}
met_log<-as.data.frame(t(met_log))
met_log_scaled<-scale(as.matrix(met_log), center = T, scale = T)
met_log_scaled_df<-as.data.frame(met_log_scaled)

```


Format df
```{r}
met_log_scaled_df$samples<-gsub("\ [1-9]*", "", rownames(met_log_scaled_df))
# add cage info
all(rownames(met_log_scaled_df) == cage$codes)
met_log_scaled_df$cage<-cage$cage

# Make values as factors
met_log_scaled_df$samples<-as.factor(met_log_scaled_df$samples)
met_log_scaled_df$cage<-as.factor(met_log_scaled_df$cage)

metabolites<-names(met_log_scaled_df)[1:67] # Subset, do not select the metadata names

# The previous analysis showed that sample "Pre Abx_OY 611" is an outlier
# Remove it
df<-met_log_scaled_df[rownames(met_log_scaled_df) != "Pre Abx_OY 611",]
# Re-order the groups factors
df$samples<-factor(df$samples, levels = c("PreAbx_YO", "PreAbx_OY", "PostFMT_OY", "PostFMT_YO"))
```

Make the graphs for each metabolite.

```{r}
met_log_scaled_df$samples<-gsub("\ [1-9]*", "", rownames(met_log_scaled_df))

# add cage info
all(rownames(met_log_scaled_df) == cage$codes)
met_log_scaled_df$cage<-cage$cage

# Make values as factors
met_log_scaled_df$samples<-as.factor(met_log_scaled_df$samples)
met_log_scaled_df$cage<-as.factor(met_log_scaled_df$cage)

p<-"Metabolome/Figures/"
dir.create(p, recursive = T)
for(i in 1:length(metabolites)){
  ggplot(met_log_scaled_df,aes(samples, met_log_scaled_df[, metabolites[i]], fill = samples)) +
  geom_boxplot(outlier.colour = "none") +
  scale_fill_brewer(palette = "Set3") +
  new_scale_fill() +
  geom_jitter(width = 0.2, size = 5, aes(fill = cage), pch = 21) +
  scale_fill_brewer(palette = "Dark2") +
  theme_bw() 
  ggsave(paste0(p, metabolites[i], ".png", sep = ""), device = "png")
}

```

# Linear models

Build linear models including cage as random effect.

```{r}
l_lm<-vector(mode = "list", length = length(metabolites))
names(l_lm)<-metabolites
for(i in 1: length(metabolites)){
  print(i)
  l_lm[[i]]<-lmer(met_log_scaled_df[,i] ~ samples  + (1|cage), data = met_log_scaled_df)
}
```


Save residuals.

```{r}
for(i in 1:length(l_lm)){ 
  path<-"Metabolome/Figures/Residuals/"
  dir.create(path, recursive = T)
  x<-paste0(path,metabolites[i], ".png", sep = "")
  trellis.device(device="png", file = x)
  print(plot(l_lm[[i]]))
  dev.off()
  png(paste0(path,metabolites[i], "_qq.png", sep = ""))
  qqnorm(residuals(l_lm[[i]]))
  qqline(residuals(l_lm[[i]]))
  dev.off()
}
```

Most of the residuals do not meet the random and normal distribution condition.
Proceed using non parametric tests.

## Non parametric tests

```{r}
library(coin)
df_met<-data.frame(metab = metabolites, pvalue = vector(length = length(metabolites)))
met_log_scaled_df$samples<-as.factor(met_log_scaled_df$samples)
for(i in 1:length(metabolites)) {
  w<-kruskal_test(met_log_scaled_df[,metabolites[i]] ~ samples | cage, data = met_log_scaled_df)
  df_met[df_met$metab == metabolites[i], 2]<-pvalue(w)
}

df_met$padjust<-p.adjust(df_met$pvalue, method = "fdr")
df_met
```

Nothing significant.


### Save Image
```{r}
save.image("Metabolome/RData/DA_metabolome.RData")

```
